---
title: "Homework 3"
subtitle: "Research Methods, Spring 2024"
author: "Taeyoung Yim"
format:
  pdf:
    output-file: "yim-t-hwk4-1"
    output-ext:  "pdf"
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---

```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra, stringr, data.table, gdata)
```

```{r}
#| include: false
#| eval: true
final.data
```

\newpage
# Summarize the Data
\vspace{.2in}
\noindent 1. Remove all SNPs, 800-series plans, and prescription drug only plans (i.e., plans that do not offer Part C benefits). Provide a box and whisker plot showing the distribution of plan counts by county over time. Do you think that the number of plans is sufficient, too few, or too many?<br>
```{r,  echo = FALSE}
library(dplyr)
library(ggplot2)

# Filter out SNPs, 800-series plans, and prescription drug only plans
filtered_data <- final.data %>%
  filter(snp == "No", !grepl("^8", planid), partd == "No")

# Summarize the number of plans by county over time
plan_counts <- filtered_data %>%
  group_by(county, year) %>%
  summarise(plan_count = n_distinct(planid))

# Box and whisker plot
ggplot(plan_counts, aes(x = year, y = plan_count)) +
  geom_boxplot() +
  labs(x = "Year", y = "Plan Count") +
  ggtitle("Distribution of Plan Counts by County Over Time")
```

\newpage
\vspace{.2in}
\noindent 2. Provide bar graphs showing the distribution of star ratings in 2010, 2012, and 2015. How has this distribution changed over time?<br>
```{r,  echo = FALSE}
library(dplyr)
library(ggplot2)

# Filter the data for the years 2010, 2012, and 2015
filtered_data <- final.data %>%
  filter(year %in% c(2010, 2012, 2015))

# Count the occurrences of each star rating for each year
star_ratings_count <- filtered_data %>%
  group_by(year, mean_risk) %>%
  summarise(count = n())

# Create bar graphs for each year
ggplot(star_ratings_count, aes(x = factor(mean_risk), y = count)) +
  geom_col() +
  facet_wrap(~year, scales = "free") +
  labs(x = "Star Rating", y = "Count") +
  ggtitle("Distribution of Star Ratings in 2010, 2012, and 2015") +
  theme_minimal()
```

\newpage
\vspace{.2in}
\noindent 3.Plot the average benchmark payment over time from 2010 through 2015. How much has the average benchmark payment risen over the years?.<br>
```{r,  echo = FALSE}
library(dplyr)
library(ggplot2)

# Filter the data for the years 2010 through 2015
filtered_data <- final.data %>%
  filter(year >= 2010 & year <= 2015)

# Calculate the average benchmark payment for each year
average_benchmark_payment <- filtered_data %>%
  group_by(year) %>%
  summarise(avg_benchmark_payment = mean(avg_ffscost, na.rm = TRUE))

# Plot the average benchmark payment over time
ggplot(average_benchmark_payment, aes(x = year, y = avg_benchmark_payment)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", y = "Average Benchmark Payment") +
  ggtitle("Average Benchmark Payment Over Time (2010 - 2015)") +
  theme_minimal()

```

\newpage
\vspace{.2in}
\noindent 4.Plot the average share of Medicare Advantage (relative to all Medicare eligibles) over time from 2010 through 2015. Has Medicare Advantage increased or decreased in popularity? How does this share correlate with benchmark payments?<br>
```{r,  echo = FALSE}
library(dplyr)
library(ggplot2)

# Filter the data for the years 2010 through 2015
filtered_data <- final.data %>%
  filter(year >= 2010 & year <= 2015)

# Calculate the total number of Medicare eligibles for each year
medicare_eligibles <- filtered_data %>%
  group_by(year) %>%
  summarise(total_medicare_eligibles = sum(avg_enrollment, na.rm = TRUE))

# Calculate the total enrollment in Medicare Advantage plans for each year
medicare_advantage_enrollment <- filtered_data %>%
  filter(partc_enroll > 0) %>%
  group_by(year) %>%
  summarise(total_medicare_advantage_enrollment = sum(partc_enroll))

# Merge the two datasets
merged_data <- merge(medicare_eligibles, medicare_advantage_enrollment, by = "year", all = TRUE)

# Calculate the average share of Medicare Advantage relative to all Medicare eligibles for each year
merged_data <- merged_data %>%
  mutate(average_share_medicare_advantage = total_medicare_advantage_enrollment / total_medicare_eligibles)

# Plot the average share of Medicare Advantage over time
ggplot(merged_data, aes(x = year, y = average_share_medicare_advantage)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", y = "Average Share of Medicare Advantage",
       title = "Average Share of Medicare Advantage Over Time (2010 - 2015)") +
  theme_minimal()

# Correlation with benchmark payments
correlation <- cor(merged_data$average_share_medicare_advantage, merged_data$total_medicare_advantage_enrollment)
correlation

```

\newpage
# Estimate ATEs
For the rest of the assignment, we’ll use a regression discontinuity design to estimate the average treatment effect from receiving a marginally higher rating. We’ll focus only on 2010.
\vspace{.2in}
\noindent 5.Calculate the running variable underlying the star rating. Provide a table showing the number of plans that are rounded up into a 3-star, 3.5-star, 4-star, 4.5-star, and 5-star rating.<br>
```{r, echo = FALSE}

```
\newpage

\vspace{.2in}
\noindent 6.Using the RD estimator with a bandwidth of 0.125, provide an estimate of the effect of receiving a 3-star versus a 2.5 star rating on enrollments. Repeat the exercise to estimate the effects at 3.5 stars, and summarize your results in a table..<br>

```{r, echo=FALSE}

```

The coefficient estimate of -0.809 for ln_price_2012 (p < 0.001) indicates that there is a substantial negative correlation between cigarette sales and pricing. This means that reduced sales correlated to increased cigarette prices. 

\newpage
\vspace{.2in}
\noindent 7.Repeat your results for bandwidhts of 0.1, 0.12, 0.13, 0.14, and 0.15 (again for 3 and 3.5 stars). Show all of the results in a graph. How sensitive are your findings to the choice of bandwidth?<br>

```{r, echo=FALSE}

```


\newpage
\vspace{.2in}
\noindent 8.Examine (graphically) whether contracts appear to manipulate the running variable. In other words, look at the distribution of the running variable before and after the relevent threshold values. What do you find?.<br>
```{r,  echo = FALSE}


```

\newpage
\vspace{.2in}
\noindent 9.Similar to question 4, examine whether plans just above the threshold values have different characteristics than contracts just below the threshold values. Use HMO and Part D status as your plan characteristics.<br>
```{r,  echo = FALSE}


```

\newpage
\vspace{.2in}
\noindent 10.Summarize your findings from 5-9. What is the effect of increasing a star rating on enrollments? Briefly explain your results.<br>
```{r,  echo = FALSE}

```