---
title: "Homework 4"
subtitle: "Research Methods, Spring 2024"
author: "Taeyoung Yim"
format:
  pdf:
    output-file: "yim-t-hwk4-1"
    output-ext:  "pdf"
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---

```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra, stringr, data.table, gdata)
```

```{r}
#| include: false
#| eval: true

final.data <- read_rds("data/output/final_ma_data.rds")
```

\newpage
# Summarize the Data
\vspace{.2in}
\noindent 1. Remove all SNPs, 800-series plans, and prescription drug only plans (i.e., plans that do not offer Part C benefits). Provide a box and whisker plot showing the distribution of plan counts by county over time. Do you think that the number of plans is sufficient, too few, or too many?<br>
```{r,  echo = FALSE}
library(ggplot2)
filtered_data <- final.data %>%
  filter(!(plan_type == "SNP" | grepl("^800", planid) | partd))

plan_counts <- filtered_data %>%
  group_by(county, year) %>%
  summarise(plan_count = n())

boxplot_plot <- ggplot(plan_counts, aes(x = as.factor(year), y = plan_count)) +
  geom_boxplot(fill = "#4C72B0", color = "#4C72B0",alpha = 0.8,outlier.shape = NA) +  
  labs(x = "Year", y = "Plan Count", title = "Distribution of Plan Counts by County Over Time") +
  theme_minimal() +ylim(0, 40)


print(boxplot_plot)
```

\newpage
\vspace{.2in}
\noindent 2. Provide bar graphs showing the distribution of star ratings in 2010, 2012, and 2015. How has this distribution changed over time?<br>
```{r, echo = FASLE}
# Define shades of blue for each year
custom_colors <- c("skyblue", "pink", "darkgreen")

# Filter the data for the years 2010, 2012, and 2015
filtered_data <- final.data[final.data$year %in% c(2010, 2012, 2015), ]

# Create bar graphs for each year
par(mfrow = c(1, 3)) # Set up a 1x3 layout for the plots
for (i in 1:3) {
  # Subset data for the current year
  year_data <- filtered_data[filtered_data$year == c(2010, 2012, 2015)[i], ]
  # Count the frequency of each star rating
  star_counts <- table(year_data$Star_Rating)
  # Create a bar plot with shades of blue for each year
  barplot(star_counts, main = paste("Star Ratings Distribution -", c(2010, 2012, 2015)[i]),
          xlab = "Star Rating", ylab = "Frequency", col = custom_colors[i])
}

```
The mean start ratings seems to increase over time. 

\newpage
\vspace{.2in}
\noindent 3.Plot the average benchmark payment over time from 2010 through 2015. How much has the average benchmark payment risen over the years?.<br>
```{r,  echo = FALSE}
filtered_data <- final.data[final.data$year %in% 2010:2015, ]

average_payment <- aggregate(payment_partc ~ year, data = filtered_data, FUN = mean)

plot(average_payment$year, average_payment$payment_partc, type = "o",
     main = "Average Benchmark Payment Over Time (2010-2015)",
     xlab = "Year", ylab = "Average Benchmark Payment")
grid()

rise <- diff(average_payment$payment_partc)
rise_over_years <- sum(rise)
cat("The average benchmark payment has risen by", round(rise_over_years, 2), "over the years (2010-2015).")
```


\newpage
\vspace{.2in}
\noindent 4.Plot the average share of Medicare Advantage (relative to all Medicare eligibles) over time from 2010 through 2015. Has Medicare Advantage increased or decreased in popularity? How does this share correlate with benchmark payments?<br>
```{r,  echo = FALSE}
# Filter the data for the years 2010 through 2015
filtered_data <- final.data[final.data$year %in% 2010:2015, ]

# Calculate the average share of Medicare Advantage enrollees relative to all Medicare eligibles for each year
average_ma_share <- filtered_data %>% group_by(fips, year)%>%
  summarize(enroll = first(avg_enrolled), medicare = first(avg_eligibles), bench = mean(ma_rate, na.rm=TRUE))%>%
  mutate(mkt_share = enroll/medicare)

ma.share <- average_ma_share%>%
  ggplot(aes(x = as.factor(year), y = mkt_share, group = 1))+
  stat_summary(fun = "mean", geom= "line", na.rm=TRUE) +
  labs(
    x = "Year",
    y = "MA Market Share"
  ) + theme_bw()
ma.share
```

Medical advantage increased overtime. Comparing this trend with the benchmark payment trend, the over all trend seems to be positively correlated, although the trend is slight different between 2011 and 2013.

\newpage
# Estimate ATEs
For the rest of the assignment, we’ll use a regression discontinuity design to estimate the average treatment effect from receiving a marginally higher rating. We’ll focus only on 2010.
\vspace{.2in}
\noindent 5.Calculate the running variable underlying the star rating. Provide a table showing the number of plans that are rounded up into a 3-star, 3.5-star, 4-star, 4.5-star, and 5-star rating.<br>
```{r, echo = FALSE}
library(rdrobust)
# Filter final.data for the year 2010
data_2010 <- final.data %>%
  filter(year == 2010)

# Calculate the running variable and associated variables for 2010 data
running_variable_2010 <- data_2010 %>%
  mutate(
    score = Star_Rating - 2.25,
    treat = (score >= 0),
    window1 = (score >= -0.175 & score <= 0.175),
    window2 = (score >= -0.125 & score <= 0.125),
    mkt_share = avg_enrollment / avg_eligibles,
    ln_share = log(mkt_share),
    score_treat = score * treat
  )

# Print summary statistics or other analyses on the calculated variables
summary(running_variable_2010)

```

\newpage

\vspace{.2in}
\noindent 6.Using the RD estimator with a bandwidth of 0.125, provide an estimate of the effect of receiving a 3-star versus a 2.5 star rating on enrollments. Repeat the exercise to estimate the effects at 3.5 stars, and summarize your results in a table.<br>

```{r, echo=FALSE}
star30 <- lm(mkt_share ~ treat + score + treat_score,
  data = (ma.data.clean %>%
    filter(raw_ratings >= (2.75-0.125)
      raw_rating <= (2.75 + 0.125)
      Star_Rating %in% c(2.5,3.0))%>%
    mutate(treat = (Star_Rating == 3.0),
      score = raw_rating - 2.75,
      treat_score = treat*score)))
```


\newpage
\vspace{.2in}
\noindent 7.Repeat your results for bandwidhts of 0.1, 0.12, 0.13, 0.14, and 0.15 (again for 3 and 3.5 stars). Show all of the results in a graph. How sensitive are your findings to the choice of bandwidth?<br>

```{r, echo=FALSE}
# Function to estimate treatment effect using linear regression for a given bandwidth
estimate_treatment_effect_regression <- function(star_rating, bandwidth) {
  # Filter data around the star rating cutoff
  data_around_cutoff <- final.data %>%
    filter(Star_Rating >= (star_rating - bandwidth) & Star_Rating <= (star_rating + bandwidth))
  
  # Fit linear regression model
  model <- lm(avg_enrolled ~ Star_Rating, data = data_around_cutoff)
  
  # Extract and return treatment effect estimate
  return(coef(model)[2])
}

# Define star ratings of interest
star_ratings <- c(3, 3.5)

# Define bandwidths of interest
bandwidths <- c(0.1, 0.12, 0.13, 0.14, 0.15)

# Initialize empty lists to store treatment effects
treatment_effects_3 <- list()
treatment_effects_3.5 <- list()

# Estimate treatment effects for each bandwidth and star rating
for (bandwidth in bandwidths) {
  treatment_effects_3[[as.character(bandwidth)]] <- estimate_treatment_effect_regression(3, bandwidth)
  treatment_effects_3.5[[as.character(bandwidth)]] <- estimate_treatment_effect_regression(3.5, bandwidth)
}

# Create a data frame for plotting
results_df_regression <- data.frame(
  Bandwidth = rep(bandwidths, each = 2),
  Star_Rating = rep(c(3, 3.5), times = length(bandwidths)),
  Treatment_Effect = c(unlist(treatment_effects_3), unlist(treatment_effects_3.5))
)

# Print the results
print(results_df_regression)

```
I am getting "null" for all, so it is not possible to tell. 

\newpage
\vspace{.2in}
\noindent 8.Examine (graphically) whether contracts appear to manipulate the running variable. In other words, look at the distribution of the running variable before and after the relevent threshold values. What do you find?.<br>
```{r,  echo = FALSE}
# Load necessary libraries
library(ggplot2)

# Define the relevant threshold values (e.g., 2.5, 3, 3.5)
threshold_values <- c(2.5, 3, 3.5)

# Initialize an empty list to store plots
plots <- list()

# Loop through each threshold value
for (threshold_value in threshold_values) {
  # Filter the data for the year 2010 and around the threshold value
  data_around_threshold <- final.data %>%
    filter(year == 2010, Star_Rating >= (threshold_value - 0.5) & Star_Rating <= (threshold_value + 0.5))
  
  # Create a density plot of the running variable before and after the threshold
  plot <- ggplot(data_around_threshold, aes(x = Star_Rating)) +
    geom_density(fill = "skyblue", alpha = 0.5) +
    geom_vline(xintercept = threshold_value, linetype = "dashed", color = "red") +
    labs(title = paste("Density Plot of Running Variable around Threshold =", threshold_value),
         x = "Star Rating", y = "Density") +
    theme_minimal()
  
  # Store the plot
  plots[[as.character(threshold_value)]] <- plot
}

# Print the plots
for (threshold_value in threshold_values) {
  print(plots[[as.character(threshold_value)]])
}


```
It peaks at 3, 3.5, and 4.
\newpage
\vspace{.2in}
\noindent 9.Similar to question 8, examine whether plans just above the threshold values have different characteristics than contracts just below the threshold values. Use HMO and Part D status as your plan characteristics.<br>
```{r,  echo = FALSE}
# Load necessary libraries
library(ggplot2)

# Define the relevant threshold values (e.g., 2.5, 3, 3.5)
threshold_values <- c(2.5, 3, 3.5)

# Initialize an empty list to store plots
plots <- list()

# Loop through each threshold value
for (threshold_value in threshold_values) {
  # Filter the data for the year 2010 and around the threshold value
  data_above_threshold <- final.data %>%
    filter(year == 2010, Star_Rating > threshold_value)
  
  data_below_threshold <- final.data %>%
    filter(year == 2010, Star_Rating < threshold_value)
  
  # Create density plots for org_type (assumed to represent HMO status)
  plot_hmo <- ggplot() +
    geom_density(data = data_above_threshold, aes(x = org_type, fill = "Above Threshold"), alpha = 0.5) +
    geom_density(data = data_below_threshold, aes(x = org_type, fill = "Below Threshold"), alpha = 0.5) +
    labs(title = paste("Density Plot of HMO Status around Threshold =", threshold_value),
         x = "HMO Status", y = "Density") +
    theme_minimal() +
    scale_fill_manual(values = c("Above Threshold" = "blue", "Below Threshold" = "red"))
  
  # Create density plots for partd (Part D status)
  plot_partd <- ggplot() +
    geom_density(data = data_above_threshold, aes(x = partd, fill = "Above Threshold"), alpha = 0.5) +
    geom_density(data = data_below_threshold, aes(x = partd, fill = "Below Threshold"), alpha = 0.5) +
    labs(title = paste("Density Plot of Part D Status around Threshold =", threshold_value),
         x = "Part D Status", y = "Density") +
    theme_minimal() +
    scale_fill_manual(values = c("Above Threshold" = "blue", "Below Threshold" = "red"))
  
  # Store the plots
  plots[[paste("HMO", as.character(threshold_value))]] <- plot_hmo
  plots[[paste("PartD", as.character(threshold_value))]] <- plot_partd
}

# Print the plots
for (plot_name in names(plots)) {
  print(plots[[plot_name]])
}

```

\newpage
\vspace{.2in}
\noindent 10.Summarize your findings from 5-9. What is the effect of increasing a star rating on enrollments? Briefly explain your results.<br>

Number 5, 6 and 7 clearly seems to be wrong at this point, so it is hard to clearly tell what is happening. My guess is that it should have positive impact on enrollment rate.